Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2014-12-16 doc/GMP2015/s04_recombination.tex
Added figures to intro. Updated tikz figs, put them in separate files.
Completed DRP example 2. Assembled recomb image. Got different images for intro and cleaned up tex file.

\section{Using the canonical, optimal DR-plan for Realizing (Solving) Qusecs}
\label{sec:recomb}
In this section, we consider the {\em optimal recombination}
problem of combining specific solutions of
subsystems in a DR-plan into a solution of their parent system $I$, (without
loss of generality, at the top level of the DR-plan). In the case of isostatic
qusecs, the parent system $I$ is isostatic (the root of the DR-plan), and we
seek  {\em solution(s) (among a finite large number of solutions)
with a specific orientation or chirality}.
In the case of underconstrained qusecs the subsystems are the multiple roots
of the DR-plan, the parent system $I$ is underconstrained, and we typically seek
an efficient algorithmic description of {\em connected component(s) of solutions with a specific orientation or
chirality}.

@@

\medskip\noindent{\bf Example}.
A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial isostatic
graphs, i.e, its DR-plan has a root and 9
children corresponding to the 9 edges. Solving or recombining the system
$(G,\delta)$
corresponding to the root of
this DR-plan involves solving a quadratic system with 8 equations and variables,
Instead of simultaneously solving this system, we could instead use the fact
that $G=K_{3,3}$ is in OMD$_1$: remove the
edge $e$ in Figure \ref{omdexample} to give a partial 2-tree $G'$.
Now add the nonedge $f$ to give a 2-tree $H$ with a DR-plan of size 2.
The Cayley configuration space $\Phi_f(G', \delta_{E\setminus e})$
is a single interval of attainable length values $\lambda_F$ for
the edge $f$.
When $\delta$ is generic, i.e, does not admit collinearities or coincidences
in the realizations of $(G,\delta)$,
the realization space of $(H, <\delta_{E\setminus e}, \lambda_f>)$ has 16
solutions $q_p^{\lambda_f}$ (modulo orientation preserving isometries),
with distinct orientation types $\sigma_p$,
(two orientation choices for each of the 4 triangles)
that can be obtained by solving a sequence of 4 single quadratics in 1
variable (DR-plan of size 2).
By subdivided
binary search
in the interval $\lambda_f \in
\Phi_f(G', \delta_{E\setminus e})$, the desired
solution $p$  of $(G,\delta)$ is found when the length of the nonedge
$e$ in  the realization
$q^p_{\lambda_f}$
is $\delta_e$.
%
\subsection{Optimized Modification by Enlarging the class of Reduced
Graphs}
\label{tdecomp}
It is possible in principle to decrease $k$ for a OMD$_k$ graph
(i.e, the number of edges to be removed
to ensure an isostatic completion that is decomposable with a small DR-plan) by
considering reduced graphs $G'$ (and modified graphs $H$) that come from a larger class than
partial 2-trees but nevertheless have convex Cayley configuration spaces at
least when the realization space is restricted to a sufficiently
comprehensive orientation type.
In particular, the so-called {\em tree-decomposable graphs of low Cayley
complexity} \cite{XX} include the partial 2-trees and many others that are
not partial 2-trees. These too result in DR-plans of size 2 or 3, putting $G$ in the class
OMD$_k$ and thus meeting Criteria (a) and (b). The Criterion (c) is met
- for example when $k=1$ -  because 1 dof Cayley configuration spaces of
linkages based on such graphs $G'$
are known to be single intervals
when a comprehensive orientation type $\sigma_p$ of the sought solution $p$ is given.
In addition, the bounds of these intervals are of low algebraic
complexity.
More precisely, the bounds  can themselves be computed using a DR-plan of size 2 or 3,
i.e, the computation of these bounds is tree-decomposable. Alternatively,
the bounds are in a simple quadratic or radically solvable extension field over the rationals,
or can be computed by solving a
triangularized system of quadratics.

>>>>

\section{Using the canonical, optimal DR-plan for Realizing (Solving) Qusecs}
\label{sec:recomb}

In this section, we consider the {\em optimal recombination}
problem of combining specific solutions of
subsystems in a DR-plan into a solution of their parent system $I$, (without
loss of generality, at the top level of the DR-plan). In the case of isostatic
qusecs, the parent system $I$ is isostatic (the root of the DR-plan), and we
seek  {\em solution(s) (among a finite large number of solutions)
with a specific orientation or chirality}.
In the case of underconstrained qusecs the subsystems are the multiple roots
of the DR-plan, the parent system $I$ is underconstrained, and we typically seek
an efficient algorithmic description of {\em connected component(s) of solutions with a specific orientation or
chirality}.

@@

\begin{figure*}\centering
\begin{subfigure}{.2\linewidth}\centering
    \input{img/tikz/omd_k33_example_new-0.tikz}
    \caption{}\label{fig:omd_k33_example:a}
\end{subfigure}\hfill
\begin{subfigure}{.15\linewidth}\centering
    \input{img/tikz/omd_k33_example_new-1.tikz}
    \caption{}\label{fig:omd_k33_example:b}
\end{subfigure}
\begin{subfigure}{.15\linewidth}\centering
    \input{img/tikz/omd_k33_example_new-2.tikz}
    \caption{}\label{fig:omd_k33_example:c}
\end{subfigure}\hfill
\begin{subfigure}{.15\linewidth}\centering
    \input{img/tikz/omd_k33_example-1.tikz}
    \caption{}\label{fig:omd_k33_example:d}
\end{subfigure}
\begin{subfigure}{.15\linewidth}\centering
    \input{img/tikz/omd_k33_example-2.tikz}
    \caption{}\label{fig:omd_k33_example:e}
\end{subfigure}

\caption{
(\ref{fig:omd_k33_example:a}) The $K3,3$ with two labeled edges, $e_1$ and $e_2$. (\ref{fig:omd_k33_example:b}) The $K3,3$ with $e_1$ and $e_2$ removed (dashed lines) and rearranged to illustrate that it is now a partial 2-tree. (\ref{fig:omd_k33_example:c}) The $K3,3$ with $\{e_1,e_2\}$ removed and $\{f_1,f_2\}$ (bold lines) added to make a 2-tree, showing that the $K3,3$ is at least $OMD_2$. (\ref{fig:omd_k33_example:d}) The $K3,3$ with only $e_2$ removed (dashed line). (\ref{fig:omd_k33_example:e}) The $K3,3$ with $e_2$ removed and $f_3$ (bold line) added to make a low Cayley complexity graph, showing that the $K3,3$ is $OMD_1$.
}
\label{fig:omd_k33_example}
\end{figure*}

\medskip\noindent{\bf Example}.
A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial isostatic
graphs, i.e, its DR-plan has a root and 9
children corresponding to the 9 edges. Solving or recombining the system
$(G,\delta)$
corresponding to the root of
this DR-plan involves solving a quadratic system with 8 equations and variables,
Instead of simultaneously solving this system, we could instead use the fact
that $G=K_{3,3}$ is in OMD$_1$: remove the
edge $e$ in Figure \ref{fig:omd_k33_example} to give a partial 2-tree $G'$.
Now add the nonedge $f$ to give a 2-tree $H$ with a DR-plan of size 2.
The Cayley configuration space $\Phi_f(G', \delta_{E\setminus e})$
is a single interval of attainable length values $\lambda_F$ for
the edge $f$.
When $\delta$ is generic, i.e, does not admit collinearities or coincidences
in the realizations of $(G,\delta)$,
the realization space of $(H, <\delta_{E\setminus e}, \lambda_f>)$ has 16
solutions $q_p^{\lambda_f}$ (modulo orientation preserving isometries),
with distinct orientation types $\sigma_p$,
(two orientation choices for each of the 4 triangles)
that can be obtained by solving a sequence of 4 single quadratics in 1
variable (DR-plan of size 2).
By subdivided
binary search
in the interval $\lambda_f \in
\Phi_f(G', \delta_{E\setminus e})$, the desired
solution $p$  of $(G,\delta)$ is found when the length of the nonedge
$e$ in  the realization
$q^p_{\lambda_f}$
is $\delta_e$.
%
\subsection{Optimized Modification by Enlarging the class of Reduced
Graphs}
\label{tdecomp}
It is possible in principle to decrease $k$ for a OMD$_k$ graph
(i.e, the number of edges to be removed
to ensure an isostatic completion that is decomposable with a small DR-plan) by
considering reduced graphs $G'$ (and modified graphs $H$) that come from a larger class than
partial 2-trees but nevertheless have convex Cayley configuration spaces at
least when the realization space is restricted to a sufficiently
comprehensive orientation type.
In particular, the so-called {\em tree-decomposable graphs of low Cayley
complexity} \cite{XX} include the partial 2-trees and many others that are
not partial 2-trees. These too result in DR-plans of size 2 or 3, putting $G$ in the class
OMD$_k$ and thus meeting Criteria (a) and (b). The Criterion (c) is met
- for example when $k=1$ -  because 1 dof Cayley configuration spaces of
linkages based on such graphs $G'$
are known to be single intervals
when a comprehensive orientation type $\sigma_p$ of the sought solution $p$ is given.
In addition, the bounds of these intervals are of low algebraic
complexity.
More precisely, the bounds  can themselves be computed using a DR-plan of size 2 or 3,
i.e, the computation of these bounds is tree-decomposable. Alternatively,
the bounds are in a simple quadratic or radically solvable extension field over the rationals,
or can be computed by solving a
triangularized system of quadratics.
